---
title: "Automated Tables and Figures"
author: Neil Shephard
date: 2017-11-14
---

# The Problem

- Disconnect between Statistical Software and submitted manuscripts.
- MHRA advised that evidence of accuracy of numbers in manuscripts required.
- Proposed solution by CTRU is yet another Quality Control check in form of "Quality Gate" form.
- First time using it on DiPEP Kim Horspool checked and signed.
- Reviewers comments highlighted some problems where wrong numbers had been copied into a table.
- **NOT** Kim's fault, mistakes happen (we're all human) but shows shortcoming of proposed "Quality Gate".

# Lessening Workload and Improving Standards

- Propose all tables and figures produced by trial Statistician.
- Produce tables in M$-Word documents for copying and pasting into manuscript.
- Produce high-quality stand-alone images for inclusion.

# Proposal

- Statisticians write SAP with sample table structures and output.
- Produce Statistical Report
- Place holders for tables and figures inserted into draft manuscripts with description of what table should contain.
- Statisticians write code to produce the specified output for copying nad pasting.

# Example - Height data

```{r sample_data_plot, echo = FALSE, results = 'asis', error = FALSE, message = FALSE, eval = TRUE}
## Seed so that this is reproducible
set.seed(87965163)
## Generate some random data on height
data <- rnorm(n = 268, mean = 130, sd = 28) %>%
        as.tibble() %>%
        mutate(row = rownames(.),
               id  = case_when(nchar(row) == 1 ~ paste0('ID', '00', row),
                               nchar(row) == 2 ~ paste0('ID', '0', row),
                               nchar(row) == 3 ~ paste0('ID', row)),
               random = runif(n = nrow(.)),
               status = ifelse(random > 0.5,
                               yes = 'Case',
                               no  = 'Control')) %>%
    dplyr::select(-row, -random)
names(data) <- gsub('value', 'height', names(data))
## Introduce some erroneous values
data <- data %>%
        mutate(height = ifelse(id == 'ID050',
                               yes = 1.236149,
                               no  = height),
               height = ifelse(id == 'ID184',
                               yes = 389.4675,
                               no  = height),
               height = ifelse(id == 'ID243',
                               yes = 491.3786,
                               no  = height))

## Plot the data to show the erroneous data points
ggplot(data, aes(height, fill = status)) +
  geom_histogram() +
  facet_wrap(~status) +
  ggtitle('Height in Cases and Controls') +
  xlab('Height (cm)') + ylab('N') +
  guides(fill = FALSE) +
  theme_bw()
```

# Example - Height data (*cont.*)

```{r sample_data_table, echo = FALSE, results = 'asis', error = FALSE, message = FALSE, eval = TRUE}
## Summarise the data
digits <- 2
summary_stats <- data %>%
                 group_by(status) %>%
                 summarise(n = n(),
                           mean = mean(height, na.rm = TRUE),
                           sd   = sd(height, na.rm = TRUE),
                           min  = min(height, na.rm = TRUE),
                           max  = max(height, na.rm = TRUE),
                           p25  = quantile(height, probs = 0.25, a.rm = TRUE),
                           p50  = quantile(height, probs = 0.50, a.rm = TRUE),
                           p75  = quantile(height, probs = 0.75, a.rm = TRUE)) %>%
                 mutate(mean_sd    = paste0(formatC(mean, digits = digits, format = "f"),
                                            " (",
                                            formatC(sd, digits = digits, format = "f"),
                                            ")"),
                        range      = paste0(formatC(min, digits = digits, format = "f"),
                                            " - ",
                                            formatC(max, digits = digits, format = "f")),
                        median_iqr = paste0(formatC(p50, digits = digits, format = "f"),
                                            " (",
                                            formatC(p25, digits = digits, format = "f"),
                                            " - ",
                                            formatC(p75, digits = digits, format = "f"),
                                            ")"))
dplyr::select(summary_stats, status, n, mean_sd, range, median_iqr) %>%
    kable(caption = 'Summary statistics for Cases and Controls',
          col.names = c('Status', 'N', 'Mean (SD)', 'Range', 'Median (IQR)'))

## Extract mean and SD for each groups
results <- list()
results$case_mean <- data %>%
                     dplyr::filter(status == 'Case') %$%
                     mean(height, na.rm = TRUE)
results$case_sd   <- data %>%
                     dplyr::filter(status == 'Case') %$%
                     sd(height, na.rm = TRUE)
results$control_mean <- data %>%
                        dplyr::filter(status == 'Control') %$%
                        mean(height, na.rm = TRUE)
results$control_sd   <- data %>%
                        dplyr::filter(status == 'Control') %$%
                        sd(height, na.rm = TRUE)


```

# Example - Height data (*cont.*)

- Clearly some erroneous data points...

```{r sample_data_erroneous, echo = FALSE, results = 'asis', error = FALSE, message = FALSE, eval = TRUE}
## Identify individuals with spurious data points
data %>%
    dplyr::filter(height < 10 | height > 250) %>%
    dplyr::select(id, status, height) %>%
    kable(caption = 'Spurious data points that need correcting')
## Write the data to file for editing
write.table(data,
            file      = 'data/original.csv',
            row.names = FALSE,
            col.names = TRUE,
            sep       = ',')

```

# Gold Standard

- Ultimately *all* manuscripts could be written in RMarkdown to avoid *any* transcription errors.
- Numbers can be included in the text directly from the software.
- This would negate the need to use "quality gate" entirely and reduce the repetitive, error prone nature of copying numbers from statistical reports to M$-Word manuscripts.
- **Don't Worry** - whilst I think this should be strived for its not going to happen, if it did correcting errors would be as simple as [this](https://www.youtube.com/watch?v=s3JldKoA0zw).
