---
title: "Automated Tables and Figures"
author: Neil Shephard
date: 2017-11-14
output: word_document
---

# Overview

This is a very simple M$-Word document that has been generated from source [RMarkdown](http://rmarkdown.rstudio.com/lesson-1.html) source files in the statistical packages [R](https://www.r-project.org/).  Its purpose is to demonstrate how using this approach mitigates the scope for errors being introduced when humans have to copy numbers from one document to another.

Initially the figures and tables are based on data that contains errors within it.

```{r sample_data_read, echo = FALSE, results = 'asis', error = FALSE, message = FALSE, eval = TRUE}
data <- read.csv('../data/original.csv')
## data <- read.csv('../data/corrected.csv')

```

## Histogram

```{r sample_data_hist, echo = FALSE, results = 'asis', error = FALSE, message = FALSE, eval = TRUE}
## Plot the data to show the erroneous data points
ggplot(data, aes(height, fill = status)) +
  geom_histogram() +
  facet_wrap(~status) +
  ggtitle('Height in Cases and Controls') +
  xlab('Height (cm)') + ylab('N') +
  guides(fill = FALSE) +
  theme_bw()

```

## Box and Whisker Plot
```{r sample_data_box, echo = FALSE, results = 'asis', error = FALSE, message = FALSE, eval = TRUE}
## Plot the data to show the erroneous data points
ggplot(data, aes(status, height, fill = status)) +
  geom_boxplot() +
  ggtitle('Height in Cases and Controls') +
  xlab('Height (cm)') + ylab('N') +
  guides(fill = FALSE) +
  theme_bw()

```

## Tabulation

```{r sample_data_table, echo = FALSE, results = 'asis', error = FALSE, message = FALSE, eval = TRUE}
## Summarise the data
digits <- 2
summary_stats <- data %>%
                 group_by(status) %>%
                 summarise(n = n(),
                           mean = mean(height, na.rm = TRUE),
                           sd   = sd(height, na.rm = TRUE),
                           min  = min(height, na.rm = TRUE),
                           max  = max(height, na.rm = TRUE),
                           p25  = quantile(height, probs = 0.25, a.rm = TRUE),
                           p50  = quantile(height, probs = 0.50, a.rm = TRUE),
                           p75  = quantile(height, probs = 0.75, a.rm = TRUE)) %>%
                 mutate(mean_sd    = paste0(formatC(mean, digits = digits, format = "f"),
                                            " (",
                                            formatC(sd, digits = digits, format = "f"),
                                            ")"),
                        range      = paste0(formatC(min, digits = digits, format = "f"),
                                            " - ",
                                            formatC(max, digits = digits, format = "f")),
                        median_iqr = paste0(formatC(p50, digits = digits, format = "f"),
                                            " (",
                                            formatC(p25, digits = digits, format = "f"),
                                            " - ",
                                            formatC(p75, digits = digits, format = "f"),
                                            ")"))
dplyr::select(summary_stats, status, n, mean_sd, range, median_iqr) %>%
    kable(caption = 'Summary statistics for Cases and Controls',
          col.names = c('Status', 'N', 'Mean (SD)', 'Range', 'Median (IQR)'))

## Extract mean and SD for each groups
results <- list()
results$case_mean <- data %>%
                     dplyr::filter(status == 'Case') %$%
                     mean(height, na.rm = TRUE)
results$case_sd   <- data %>%
                     dplyr::filter(status == 'Case') %$%
                     sd(height, na.rm = TRUE)
results$control_mean <- data %>%
                        dplyr::filter(status == 'Control') %$%
                        mean(height, na.rm = TRUE)
results$control_sd   <- data %>%
                        dplyr::filter(status == 'Control') %$%
                        sd(height, na.rm = TRUE)

```

## In-line text

Its even possible to have the text based on the underlying data updated automatically.  For example the mean in cases is `r results$case_mean` (SD `r results$case_sd`) whilst in controls the mean is  `r results$control_mean` (SD `r results$control_sd`).
